%%%%%%%%%%%%%%%%%%%%%%%%%%
% NUMERICAL CASE STUDIES %
%%%%%%%%%%%%%%%%%%%%%%%%%%
In order to illustrate the power and versatility of the devised framework we conduct a selection of computer experiments.
This shall be seen as a proof of concept and benchmark of the proposed methodology in the context of engineering applications.
% FORWARD PROBLEM: MULTIPLE BEAMS
A system of identically designed structural components functions as the basis for probing a range of experimental scenarios.
Specifically we deal with an ensemble of simply supported beams that are tested in a series of three-point bending experiments.
% INVERSE PROBLEMS UNDER UNCERTAINTY
By multilevel analysis of measured beam deflections we highlight how different inferential goals, e.g.\ probabilistic inversion, residual calibration or optimal combination of information,
can be achieved in the presence of material variability and uncertainties in the experimental setup.
% DETERMINISTIC MODELING
Keeping deterministic modeling simple and intuitive will allow us to focus on uncertainty quantification aspects that are the essential subject matter of this research.
% COMPUTATIONAL OBSTACLES
Incidentally we learn about the computational obstacles that must be overcome when aiming at ``real-world'' applications.
\par % SECTION OUTLINE
The forward problem will be shortly introduced in \cref{sec:PEM:CaseStudies:Beams}.
Around this submodel, that covers the deterministic features of the system, Bayesian multilevel models will be built to capture uncertainty and variability.
Probabilistic inversion, i.e.\ deducing the material variability throughout an ensemble of similar specimens, will be tackled in \cref{sec:PEM:CaseStudies:ProbInv}.
The subsequent \cref{sec:PEM:CaseStudies:ResCal} will deal with residual model calibration.
In \cref{sec:PEM:CaseStudies:AddPres} the impact of prescribed uncertainties in the test conditions will be investigated.
In \cref{sec:PEM:CaseStudies:CombInf} borrowing strength will be utilized in order to ideally estimate the material characteristics of a single specimen by using information obtained from the other specimens.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mechanical model} \label{sec:PEM:CaseStudies:Beams}
\input{tex/PEM/tex_PEM_8_1_MechanicalModel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Probabilistic inversion} \label{sec:PEM:CaseStudies:ProbInv}
\input{tex/PEM/tex_PEM_8_2_ProbabilisticInversion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Residual calibration} \label{sec:PEM:CaseStudies:ResCal}
\input{tex/PEM/tex_PEM_8_3_ResidualCalibration}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Uncertain conditions} \label{sec:PEM:CaseStudies:AddPres}
\input{tex/PEM/tex_PEM_8_4_AdditionalNuisance}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Borrowing strength} \label{sec:PEM:CaseStudies:CombInf}
\input{tex/PEM/tex_PEM_8_5_BorrowingStrength}