% PROOF OF CONCEPT
Next, the potential and the difficulties of the theory presented in the preceding section are investigated.
The goal is to give a proof of concept for the basic feasibility of spectral Bayesian inference.
It is verified that the theory can be successfully applied in practice and further insight into its functioning is obtained.
Moreover, it is learned about its current shortcomings.
Four instructive calibration problems from classical statistics and inverse modeling are solved for these purposes.
The analysis is confined to problems with low-dimensional parameter spaces.
First, the mean value of a normal distribution is inferred with random data under a conjugate normal prior.
Second, the mean and standard deviation of a normal distribution are fitted for a joint prior with independent and uniform marginals.
Third, an inverse heat conduction problem in two spatial dimensions with two unknowns is solved.
Finally, a similar thermal problem with six unknowns is considered.
Synthetically created pseudo-data are used in all these example applications.
\par % CONVERGENCE BEHAVIOR
As it turns out, one can gain valuable insights into the characteristics of likelihood expansions and posterior emulators by way of comparison.
Therefore, the analyses for the first three examples proceed analogously.
For rich experimental designs, the convergence behavior of high-degree SLEs is studied by reference to the LOO error.
More importantly, the capability of lower-degree SLEs to accurately capture the posterior QoI-expectations is explored for scarcer experimental designs.
Eventually, aSLE-based posterior surrogates are investigated in order to mitigate the curse of dimensionality.
All results are compared to reference solutions.
Where possible, the exact solutions from a conjugate Bayesian analysis are used to this effect.
Otherwise, corresponding approximations are computed via classical MCMC sampling.
\par % UQLAB
The uncertainty quantification platform UQLab \cite{Computing:Marelli2014:Proc,Computing:Uqlab2015:Manual09104} is used throughout the numerical demonstrations.
It provides a flexible environment for the uncertainty analysis of engineering systems, e.g.\ for uncertainty propagation.
In this context it ships with a range of regression tools that allow one to easily compute PCEs.
These tools can be directly applied to the likelihood function in order to compute SLEs.
OLS is employed as the standard solving routine in the following examples.

\subsection{1D normal fitting}
\input{tex/JCP/tex_JCP_5_1_NormalFitting1D}

\subsection{2D normal fitting}
\input{tex/JCP/tex_JCP_5_2_NormalFitting2D}

\subsection{2D inverse heat conduction}
\input{tex/JCP/tex_JCP_5_3_HeatConduction2D}

\subsection{6D inverse heat conduction}
\input{tex/JCP/tex_JCP_5_4_HeatConduction6D}